Inefficient Concessions and Mediation

### 2023-08-10 draft_narrow.tex
tweak last sentence of intro

These empirical results have until recently lacked strong theoretical foundations. Mediation is not uncommon, but we are only beginning to understand why actors use it. Recently, mediation has begun to be modeled using mechanism design for settings where militarization or the resolve to fight are private information (Bester and Warneryd 2006; Fey and Ramsay 2009, 2010 and 2011; Horner et. al. 2010; Meirowitz et al. 2012, 2019). In particular, Meirowitz et al. 2019 shows that---when asymmetric information is about the level of militarization---a `Myerson' mediator can circumvent the incentives for militarization that arises in unmediated peace talks and improve the changes for peace. This paper expands the theoretical literature by modeling the role of a manipulative mediator when the information asymmetry is about parties' abilities to commit to cooperation and concessions have future material value that could be used to harm the giver. Here, a mediator can achieve peace (1) in situations where bilateral concessions cannot, and (2) with concessions that are more efficient when bilateral concessions must be inefficient to achieve peace.

>>>>

These empirical results have until recently lacked strong theoretical foundations. Mediation is not uncommon, but we are only beginning to understand why actors use it. Recently, mediation has begun to be modeled using mechanism design for settings where militarization or the resolve to fight are private information (Bester and Warneryd 2006; Fey and Ramsay 2009, 2010 and 2011; Horner et. al. 2010; Meirowitz et al. 2012, 2019). In particular, Meirowitz et al. 2019 shows that---when asymmetric information is about the level of militarization---a `Myerson' mediator can circumvent the incentives for militarization that arises in unmediated peace talks and improve the changes for peace. This paper expands the theoretical literature by modeling the role of a manipulative mediator when the information asymmetry is about parties' abilities to commit to cooperation and concessions have future material value that could be used to harm the giver. Here, a mediator can achieve peace or the provision of other public goods (1) in situations where bilateral concessions cannot, and (2) with concessions that are more efficient when bilateral concessions must be inefficient to achieve peace.
